---
title: "Capstone - Fall, 2019"
author: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

Commented header  
=========================================

```{r echo = TRUE}
# Course: BUAN 5510
# Title: Capstone- Business Analytics
# Purpose: King's County Court House
# Date: 2019-10-10
# Author: Sourabh Gupta, Keith Castelino
```

Clear environment of variables and packages  
=========================================

```{r message=FALSE, warning=FALSE, include=FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 
# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)
```

Load packages that include different functions  
=========================================

```{r message = FALSE, warning=FALSE}
# adding library
# install.packages("stringi")
library(tidyverse)
# library(GGally)
# library(gridExtra)
library("readxl")
library(janitor)
# library(stringr)
# library(aod)
# library(stringi)
# library(ggplot2)
# library(dplyr)
```

Load data and begin new EDA  
=========================================

```{r message = FALSE, warning=FALSE}
#loading data
#
# xlsx files -- Data for SU Part 1 - Filed Cases and Hearings
D11_filed_cases <- read_excel("../Data/Data for SU Part 1 - Filed Cases and Hearings.xlsx", sheet = "FiledCases SU") %>% clean_names()
D12_case_event <- read_excel("../Data/Data for SU Part 1 - Filed Cases and Hearings.xlsx", sheet = "CaseEvents SU") %>% clean_names()
D13_case_event_key <- read_excel("../Data/Data for SU Part 1 - Filed Cases and Hearings.xlsx", sheet = "CaseEvent Key") %>% clean_names()
D14_case_type <- read_excel("../Data/Data for SU Part 1 - Filed Cases and Hearings.xlsx", sheet = "CaseType Key") %>% clean_names()
#
# xlsx files -- Data for SU Part 2
D21_charge <- read_excel("../Data/Data for SU Part 2.xlsx", sheet = "SU Charges") %>% clean_names()
D22_charge_def <- read_excel("../Data/Data for SU Part 2.xlsx", sheet = "Charge Definitions") %>% clean_names()
D23_dispo_key <- read_excel("../Data/Data for SU Part 2.xlsx", sheet = "Disposition Key") %>% clean_names()
D24_crime_hist <- read_excel("../Data/Data for SU Part 2.xlsx", sheet = "SU CrimHist") %>% clean_names()
#
# xlsx files -- Data for SU Part 3 - custody history
D31_custody_hist <- read_excel("../Data/Data for SU Part 3 - custody history.xlsx", sheet = "CustodyHistory SU") %>% clean_names()
#
```


Delete columns not required
=============================================
```{r}
###
D11_filed_cases <- D11_filed_cases %>% select(c(file_number, defendant_id, dob_anon, gender))

###
D21_charge <- D21_charge %>% select(-c(referral_date, offense_start_date, offense_end_date, 
                                       referral_charge_s, original_charge_s, warning))

```

Mutate D11 Data
=============================================
```{r}
# Calculate the age of the defendant
D11_filed_cases <- D11_filed_cases %>%  mutate(age_in_days = Sys.Date() - as.Date(D11_filed_cases$dob_anon))

# Convert age into categories as mentioned in Data dictionary
D11_filed_cases <-D11_filed_cases %>%  
                  mutate(age = age_in_days / 365,
                         age_cate = ifelse(age < 19, '<19', 
                                          ifelse(age > 19 & age < 30, '20-29',
                                                ifelse(age >= 30 & age < 40, '30-39',
                                                      ifelse(age >=40  & age < 50, '40-49',
                                                            ifelse(age >= 50 & age < 60, '50-59',
                                                                  ifelse(age >= 60, '>60',0))))))) %>%
                  select(c(file_number, defendant_id, dob_anon, gender, age_cate))


```

How many defendants in multiple age categories

```{r}
ggplot(data=D11_filed_cases, aes(x=age_cate)) +
geom_bar() +
geom_text(stat='count', aes(label=..count..), vjust=-1)
```

+ There are very less number of defendants having age less than 19 anf greater than 60


Mutate D13 Data
=============================================
```{r}
# Convert Case Events into Categories, A, B, C, D
# where A - Arraingment, B - Case setting, C - Hearing, D - Disposition

a_arraingment <- c("HRARRSC", "HRARDR")
b_case_setting <- c("HRCSGP","HRCSNWK","HRCSNWS","HRCSWK","HRCSWS")
c_hearing <- c("HRCOMP", "HRCOMPAR", "HRCOMPBS", "HRCOMPCS", "HRCOMPKA", "HRCOMPKC", "HRCOMPKT", "HRCOMPTR")
d_dispositon <- c()
dfa <- c("DEFFTA")

# mutate
D12_case_event <-D12_case_event %>%  
                  mutate(case_event_stage = ifelse(event_code %in% a_arraingment , 'A - Arraingment', 
                                                  ifelse(event_code %in% b_case_setting , 'B - Case setting',
                                                        ifelse(event_code %in% c_hearing , 'C - Hearing',
                                                              ifelse(event_code %in% d_dispositon , 'D - Disposition', 
                                                                     ifelse(event_code %in% dfa , 'DFA', event_code))))))



```

How many defendants in multiple case categories

```{r}
ggplot(data=D12_case_event, aes(x=case_event_stage)) +
geom_bar() +
geom_text(stat='count', aes(label=..count..), vjust=-1)
```


```{r}
###
# Split "Current Charge(s)" at '-' to get the charge code and merge it with D22 Charge definition table
D21_charge <- D21_charge %>%
  separate(current_charge_s, c("charge_code"), sep = "-", extra = 'drop')

# # Merge D11_filed_cases_SU and D12_case_events_SU tables on "File Number" and "Defendant ID"
# filed_case_events =  merge(x = D11_filed_cases, y = D12_case_event, by = c("file_number", "defendant_id"), all = TRUE) %>% select(-c("venue.x", "age_in_days", "age"))

##### Merge D21 and D23
D21_D23 =  merge(x = D21_charge, y = D23_dispo_key, by = "disposition_code", all = TRUE) %>%
                      select(c("file_number", "filing_date", "charge_code", "enhancements",
                                "disposition_date", "disposition_code", "description.x", "ranking", "type"))

ggplot(data=D21_D23, aes(x=type)) +
geom_bar() +
geom_text(stat='count', aes(label=..count..), vjust=-0.5)

```


```{r}
# Merge Filed_cases_SU and SU_charges tables on Filenumber: we will be having charges and disposition date per file
filed_case_charges =  merge(x = D11_filed_cases, y = D21_charge, by = "file_number", all = TRUE) %>%
                      select(-c("dob_anon", "referral_date.x", "referral_date.y", "referral_charge_s",
                                "original_charge_s", "age_in_days", "age"))

#----------
#----------
#####------Code not working, need to check--------#####
#----------

# filed_case_charges <- transform(filed_case_charges, charge_code = as.numeric(charge_code))
# 
# # Left Join filed_case_charges table with D22_charge_def to get class, violent and seriousness of charge
# filed_case_charges_degree = left_join(filed_case_charges, D22_charge_def, by = c("charge_code" = "Code"))
#----------
#----------  
  

```


Basic EDA
============================================

# ----

```{r}

```


```{r}

```

# ----

```{r}

```

# ----

```{r}

```

