---
title: "BUAN 5510 - Capstone Project - Linear Regression"
author: "Keith Castelino and Sourabh Gupta"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
	fig.width = 8.5,
	# include = FALSE,
	# echo = FALSE,
	message = FALSE,
	warning = FALSE
)

options(tibble.print_max = 40, tibble.print_min = 25)

library(tidyverse)
library(here)
library(broom)
library(caTools)
library(MLmetrics)

```

# Read in the Data

Read in the dataset.

```{r}

dispo_by_category <- read_csv(here::here("data", "processed", "dispo_by_category.csv"))
represent_charge_by_file <- read_csv(here::here("data", "processed", "represent_charge_by_file.csv"))
case_competency_hearing <- read_csv(here::here("data", "processed", "case_competency_hearing.csv"))
case_trial_set_event <- read_csv(here::here("data", "processed", "case_trial_set_event.csv"))
defendant_ages <- read_csv(here::here("data", "processed", "defendant_ages.csv"))
# defendant_seriousness <- read_csv(here::here("data", "processed", "defendant_seriousness.csv"))
defendant_drug_charge <- read_csv(here::here("data", "processed", "defendant_drug_charge.csv"))
defendant_drug_court <- read_csv(here::here("data", "processed", "defendant_drug_court.csv"))

base_data <- dispo_by_category %>% 
  # select(-c(age, age_group)) %>% 
  left_join(represent_charge_by_file, by = c("file_number" = "file_number")) %>% 
  left_join(case_competency_hearing, by = c("file_number" = "file_number")) %>% 
  left_join(case_trial_set_event, by = c("file_number" = "file_number")) %>% 
  # left_join(defendant_ages %>% select(-c(defendant_id, age_group)), by = c("file_number" = "file_number")) %>% 
  # left_join(defendant_seriousness %>% select(-defendant_id), by = c("file_number" = "file_number")) %>% 
  left_join(defendant_drug_charge, by = c("file_number" = "file_number")) %>% 
  left_join(defendant_drug_court, by = c("file_number" = "file_number")) %>% 
  replace_na(list(class.x = "Other", serious_cat = "Other", defendant_status = "Other", seriousness = 0, no_of_trial_sets = 0)) %>% 
  mutate(
    is_rep_charge = if_else(charge_code.x == charge_code | is.na(charge_code.x), 1, 0),
    class.x = if_else(class.x %in% c("A", "B", "C"), class.x, "GM"),
    class_a = if_else(class.x == "A", 1, 0),
    age_below_18 = if_else(age <= 18, 1, 0),
    age_below_60 = if_else(age >= 60, 1, 0),
    is_felony = if_else(class.x %in% c("A", "B", "C"), 1, 0),
    is_drug_charge = if_else(no_of_drug_charges > 0, 1, 0),
    is_drug_court = if_else(no_of_drug_court > 0, 1, 0),
    is_trial_set = if_else(no_of_trial_sets > 0, 1, 0)
  ) %>% 
  filter(
    !is.na(dispo_time_arraign),
    age_group != "Unknown",
    gender != "Unknown",
    class.x %in% c("M", "A", "B", "C", "GM"),
    defendant_status %in% c("IN", "OUT", "Other"),
    between(dispo_time_filing, 1, 5 * 365),
    between(dispo_time_arraign, 1, 5 * 365)
  ) %>% 
  mutate_at(c("age_group", "gender", "crime_hist", "class.x", "serious_cat", "defendant_status"), factor) %>% 
  dplyr::select(
    file_number, defendant_id,
    dispo_time_filing, dispo_time_arraign, gender, age, age_group, age_below_18, age_below_60,
    class.x, class_a, is_felony, seriousness, serious_cat, crime_hist, total_dfa,
    no_of_drug_charges, no_of_drug_court, no_of_trial_sets,
    is_drug_charge, is_drug_court, is_trial_set, is_competency_hearing,
    defendant_status
  ) %>%
  rename(charge_class = class.x)

base_data$age_group <- relevel(base_data$age_group, "LE_18")
base_data$gender <- relevel(base_data$gender, "Female")
base_data$crime_hist <- relevel(base_data$crime_hist, "No")
base_data$charge_class <- relevel(base_data$charge_class, "GM")
base_data$defendant_status <- relevel(base_data$defendant_status, "IN")

```

Split into test and train sets

```{r}
set.seed(101)
sample <- sample.split(base_data$dispo_time_filing, SplitRatio = 0.80 * nrow(base_data))
train <- subset(base_data, sample == TRUE)
test <- subset(base_data, sample == FALSE)
```

# Model Selection

We will be choosing our final model based on the following variables that we have found may have a numerical relationship with disposition time, per our exploratory analysis results.

+ Age: either ...
    + age (actual)
    + age group
+ Gender
+ Criminal history (whether there is past history)
+ Charge type and seriousness:
    + charge class (Actual)
    + charge class A or not (binary)
    + felony charge or not
    + seriousness value
    + seriousness categorization
+ Failure to appear
    + no of times defendant fails to appear
+ Events and charges:
    + trial set events: no of trial setting events, or whether there were trial setting events or not
    + competency hearing: whether there were competency hearings or not
+ Events and charges - Drug related:
    + drug court: number of drug court hearings, or whether there were drug court hearings or not (binary)
    + drug charge: number of drug related charges, or whether there were drug related charges or not (binary)

The dependent variable is the time to disposition, either by filing date or arraignment date.

There may be possible interactions between the following pairs of variables:

+ number of DFAs AND charge class
+ age group AND drug case

```{r}

# # SETS OF MODELS
# 
# age + gender + crime_hist + charge_class + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
# age + gender + crime_hist + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
# age + gender + crime_hist + serious_cat + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
# age + gender + crime_hist + class_a + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
# age + gender + crime_hist + is_felony + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
# 
# age_group + gender + crime_hist + charge_class + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
# age_group + gender + crime_hist + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
# age_group + gender + crime_hist + serious_cat + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
# age_group + gender + crime_hist + class_a + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
# age_group + gender + crime_hist + is_felony + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
# 
# age + gender + crime_hist + charge_class + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
# age + gender + crime_hist + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
# age + gender + crime_hist + serious_cat + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
# age + gender + crime_hist + class_a + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
# age + gender + crime_hist + is_felony + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
# 
# age_group + gender + crime_hist + charge_class + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
# age_group + gender + crime_hist + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
# age_group + gender + crime_hist + serious_cat + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
# age_group + gender + crime_hist + class_a + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
# age_group + gender + crime_hist + is_felony + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge

```

## Summary - Models Chosen

Based on our interpretation of the model output, we have concluded that Models 1 and 2 best explain the relationships different variables have with disposition time.

+ Those models show the highest R-squared at approx 0.30, the (approx) lowest MAE at 96 and (approx) lowest MAPE at 0.78
+ Charge class accounts for some of the variation in disposition time due to cases involving a drug charge. 
    + Hence the drug charge variable is shown to not be significant in _Model 1_
+ However, cases that include a drug charge include varying seriousness levels due to other charges being involved in the same case. 
    + Hence, the variable indicating the case has a drug charge is significant in _Model 2_, unlike in _Model 1_

## Model Selection - based on filing date

Initialize data frame to record key metrics

```{r}
eval_filing <- data.frame(model_nbr = numeric(), r_squared = numeric(), mae = numeric(), mape = numeric())
```

### Model 1

```{r}

model_1 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age + age_below_18 + age_below_60 + gender + crime_hist + charge_class + total_dfa + is_trial_set + is_competency_hearing + is_drug_court + is_drug_charge
)
summary(model_1)
result_1 <- tidy(model_1)
y_pred_1 <- predict.lm(model_1, test)
y_true_1 <- test$dispo_time_filing

r_sq_1 <- round(summary(model_1)$r.squared, 4)
mae_1 <- round(MLmetrics::MAE(y_pred_1, y_true_1), 4)
mape_1 <- round(MLmetrics::MAPE(y_pred_1, y_true_1), 4)

paste("R-Squared:", r_sq_1)
paste("MAE - Mean Absolute Error:", mae_1)
paste("MAPE - Mean Absolute Percentage Error:", mape_1)

eval_filing[1,] <- c(1, r_sq_1, mae_1, mape_1)

```

### Model 2

```{r}

model_2 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age + gender + crime_hist + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court + is_drug_charge
)
summary(model_2)
result_2 <- tidy(model_2)
y_pred_2 <- predict.lm(model_2, test)
y_true_2 <- test$dispo_time_filing

r_sq_2 <- round(summary(model_2)$r.squared, 4)
mae_2 <- round(MLmetrics::MAE(y_pred_2, y_true_2), 4)
mape_2 <- round(MLmetrics::MAPE(y_pred_2, y_true_2), 4)

paste("R-Squared:", r_sq_2)
paste("MAE - Mean Absolute Error:", mae_2)
paste("MAPE - Mean Absolute Percentage Error:", mape_2)

eval_filing[2,] <- c(2, r_sq_2, mae_2, mape_2)

```

### Model 3

```{r}

model_3 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age + age_below_18 + age_below_60 + gender + crime_hist + serious_cat + total_dfa + is_trial_set + is_competency_hearing + is_drug_court + is_drug_charge
)
summary(model_3)
result_3 <- tidy(model_3)
y_pred_3 <- predict.lm(model_3, test)
y_true_3 <- test$dispo_time_filing

r_sq_3 <- round(summary(model_3)$r.squared, 4)
mae_3 <- round(MLmetrics::MAE(y_pred_3, y_true_3), 4)
mape_3 <- round(MLmetrics::MAPE(y_pred_3, y_true_3), 4)

paste("R-Squared:", r_sq_3)
paste("MAE - Mean Absolute Error:", mae_3)
paste("MAPE - Mean Absolute Percentage Error:", mape_3)

eval_filing[3,] <- c(3, r_sq_3, mae_3, mape_3)

```

### Model 4

```{r}

model_4 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age + gender + crime_hist + class_a + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_4)
result_4 <- tidy(model_4)
y_pred_4 <- predict.lm(model_4, test)
y_true_4 <- test$dispo_time_filing

r_sq_4 <- round(summary(model_4)$r.squared, 4)
mae_4 <- round(MLmetrics::MAE(y_pred_4, y_true_4), 4)
mape_4 <- round(MLmetrics::MAPE(y_pred_4, y_true_4), 4)

paste("R-Squared:", r_sq_4)
paste("MAE - Mean Absolute Error:", mae_4)
paste("MAPE - Mean Absolute Percentage Error:", mape_4)

eval_filing[4,] <- c(4, r_sq_4, mae_4, mape_4)

```

### Model 5

```{r}

model_5 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age + gender + crime_hist + is_felony + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_5)
result_5 <- tidy(model_5)
y_pred_5 <- predict.lm(model_5, test)
y_true_5 <- test$dispo_time_filing

r_sq_5 <- round(summary(model_5)$r.squared, 4)
mae_5 <- round(MLmetrics::MAE(y_pred_5, y_true_5), 4)
mape_5 <- round(MLmetrics::MAPE(y_pred_5, y_true_5), 4)

paste("R-Squared:", r_sq_5)
paste("MAE - Mean Absolute Error:", mae_5)
paste("MAPE - Mean Absolute Percentage Error:", mape_5)

eval_filing[5,] <- c(5, r_sq_5, mae_5, mape_5)

```

### Model 6

```{r}

model_6 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age_group + gender + crime_hist + charge_class + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_6)
result_6 <- tidy(model_6)
y_pred_6 <- predict.lm(model_6, test)
y_true_6 <- test$dispo_time_filing

r_sq_6 <- round(summary(model_6)$r.squared, 4)
mae_6 <- round(MLmetrics::MAE(y_pred_6, y_true_6), 4)
mape_6 <- round(MLmetrics::MAPE(y_pred_6, y_true_6), 4)

paste("R-Squared:", r_sq_6)
paste("MAE - Mean Absolute Error:", mae_6)
paste("MAPE - Mean Absolute Percentage Error:", mape_6)

eval_filing[6,] <- c(6, r_sq_6, mae_6, mape_6)

```

### Model 7

```{r}

model_7 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age_group + gender + crime_hist + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_7)
result_7 <- tidy(model_7)
y_pred_7 <- predict.lm(model_7, test)
y_true_7 <- test$dispo_time_filing

r_sq_7 <- round(summary(model_7)$r.squared, 4)
mae_7 <- round(MLmetrics::MAE(y_pred_7, y_true_7), 4)
mape_7 <- round(MLmetrics::MAPE(y_pred_7, y_true_7), 4)

paste("R-Squared:", r_sq_7)
paste("MAE - Mean Absolute Error:", mae_7)
paste("MAPE - Mean Absolute Percentage Error:", mape_7)

eval_filing[7,] <- c(7, r_sq_7, mae_7, mape_7)

```

### Model 8

```{r}

model_8 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age_group + gender + crime_hist + serious_cat + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_8)
result_8 <- tidy(model_8)
y_pred_8 <- predict.lm(model_8, test)
y_true_8 <- test$dispo_time_filing

r_sq_8 <- round(summary(model_8)$r.squared, 4)
mae_8 <- round(MLmetrics::MAE(y_pred_8, y_true_8), 4)
mape_8 <- round(MLmetrics::MAPE(y_pred_8, y_true_8), 4)

paste("R-Squared:", r_sq_8)
paste("MAE - Mean Absolute Error:", mae_8)
paste("MAPE - Mean Absolute Percentage Error:", mape_8)

eval_filing[8,] <- c(8, r_sq_8, mae_8, mape_8)

```

### Model 9

```{r}

model_9 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age_group + gender + crime_hist + class_a + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_9)
result_9 <- tidy(model_9)
y_pred_9 <- predict.lm(model_9, test)
y_true_9 <- test$dispo_time_filing

r_sq_9 <- round(summary(model_9)$r.squared, 4)
mae_9 <- round(MLmetrics::MAE(y_pred_9, y_true_9), 4)
mape_9 <- round(MLmetrics::MAPE(y_pred_9, y_true_9), 4)

paste("R-Squared:", r_sq_9)
paste("MAE - Mean Absolute Error:", mae_9)
paste("MAPE - Mean Absolute Percentage Error:", mape_9)

eval_filing[9,] <- c(9, r_sq_9, mae_9, mape_9)

```

### Model 10

```{r}

model_10 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age_group + gender + crime_hist + is_felony + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_10)
result_10 <- tidy(model_10)
y_pred_10 <- predict.lm(model_10, test)
y_true_10 <- test$dispo_time_filing

r_sq_10 <- round(summary(model_10)$r.squared, 4)
mae_10 <- round(MLmetrics::MAE(y_pred_10, y_true_10), 4)
mape_10 <- round(MLmetrics::MAPE(y_pred_10, y_true_10), 4)

paste("R-Squared:", r_sq_10)
paste("MAE - Mean Absolute Error:", mae_10)
paste("MAPE - Mean Absolute Percentage Error:", mape_10)

eval_filing[10,] <- c(10, r_sq_10, mae_10, mape_10)

```

### Model 11

```{r}

model_11 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age + gender + crime_hist + charge_class + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_11)
result_11 <- tidy(model_11)
y_pred_11 <- predict.lm(model_11, test)
y_true_11 <- test$dispo_time_filing

r_sq_11 <- round(summary(model_11)$r.squared, 4)
mae_11 <- round(MLmetrics::MAE(y_pred_11, y_true_11), 4)
mape_11 <- round(MLmetrics::MAPE(y_pred_11, y_true_11), 4)

paste("R-Squared:", r_sq_11)
paste("MAE - Mean Absolute Error:", mae_11)
paste("MAPE - Mean Absolute Percentage Error:", mape_11)

eval_filing[11,] <- c(11, r_sq_11, mae_11, mape_11)

```

### Model 12

```{r}

model_12 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age + gender + crime_hist + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_12)
result_12 <- tidy(model_12)
y_pred_12 <- predict.lm(model_12, test)
y_true_12 <- test$dispo_time_filing

r_sq_12 <- round(summary(model_12)$r.squared, 4)
mae_12 <- round(MLmetrics::MAE(y_pred_12, y_true_12), 4)
mape_12 <- round(MLmetrics::MAPE(y_pred_12, y_true_12), 4)

paste("R-Squared:", r_sq_12)
paste("MAE - Mean Absolute Error:", mae_12)
paste("MAPE - Mean Absolute Percentage Error:", mape_12)

eval_filing[12,] <- c(12, r_sq_12, mae_12, mape_12)

```

### Model 13

```{r}

model_13 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age + gender + crime_hist + serious_cat + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_13)
result_13 <- tidy(model_13)
y_pred_13 <- predict.lm(model_13, test)
y_true_13 <- test$dispo_time_filing

r_sq_13 <- round(summary(model_13)$r.squared, 4)
mae_13 <- round(MLmetrics::MAE(y_pred_13, y_true_13), 4)
mape_13 <- round(MLmetrics::MAPE(y_pred_13, y_true_13), 4)

paste("R-Squared:", r_sq_13)
paste("MAE - Mean Absolute Error:", mae_13)
paste("MAPE - Mean Absolute Percentage Error:", mape_13)

eval_filing[13,] <- c(13, r_sq_13, mae_13, mape_13)

```

### Model 14

```{r}

model_14 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age + gender + crime_hist + class_a + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_14)
result_14 <- tidy(model_14)
y_pred_14 <- predict.lm(model_14, test)
y_true_14 <- test$dispo_time_filing

r_sq_14 <- round(summary(model_14)$r.squared, 4)
mae_14 <- round(MLmetrics::MAE(y_pred_14, y_true_14), 4)
mape_14 <- round(MLmetrics::MAPE(y_pred_14, y_true_14), 4)

paste("R-Squared:", r_sq_14)
paste("MAE - Mean Absolute Error:", mae_14)
paste("MAPE - Mean Absolute Percentage Error:", mape_14)

eval_filing[14,] <- c(14, r_sq_14, mae_14, mape_14)

```

### Model 15

```{r}

model_15 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age + gender + crime_hist + is_felony + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_15)
result_15 <- tidy(model_15)
y_pred_15 <- predict.lm(model_15, test)
y_true_15 <- test$dispo_time_filing

r_sq_15 <- round(summary(model_15)$r.squared, 4)
mae_15 <- round(MLmetrics::MAE(y_pred_15, y_true_15), 4)
mape_15 <- round(MLmetrics::MAPE(y_pred_15, y_true_15), 4)

paste("R-Squared:", r_sq_15)
paste("MAE - Mean Absolute Error:", mae_15)
paste("MAPE - Mean Absolute Percentage Error:", mape_15)

eval_filing[15,] <- c(15, r_sq_15, mae_15, mape_15)

```

### Model 16

```{r}

model_16 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age_group + gender + crime_hist + charge_class + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_16)
result_16 <- tidy(model_16)
y_pred_16 <- predict.lm(model_16, test)
y_true_16 <- test$dispo_time_filing

r_sq_16 <- round(summary(model_16)$r.squared, 4)
mae_16 <- round(MLmetrics::MAE(y_pred_16, y_true_16), 4)
mape_16 <- round(MLmetrics::MAPE(y_pred_16, y_true_16), 4)

paste("R-Squared:", r_sq_16)
paste("MAE - Mean Absolute Error:", mae_16)
paste("MAPE - Mean Absolute Percentage Error:", mape_16)

eval_filing[16,] <- c(16, r_sq_16, mae_16, mape_16)

```

### Model 17

```{r}

model_17 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age_group + gender + crime_hist + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_17)
result_17 <- tidy(model_17)
y_pred_17 <- predict.lm(model_17, test)
y_true_17 <- test$dispo_time_filing

r_sq_17 <- round(summary(model_17)$r.squared, 4)
mae_17 <- round(MLmetrics::MAE(y_pred_17, y_true_17), 4)
mape_17 <- round(MLmetrics::MAPE(y_pred_17, y_true_17), 4)

paste("R-Squared:", r_sq_17)
paste("MAE - Mean Absolute Error:", mae_17)
paste("MAPE - Mean Absolute Percentage Error:", mape_17)

eval_filing[17,] <- c(17, r_sq_17, mae_17, mape_17)

```

### Model 18

```{r}

model_18 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age_group + gender + crime_hist + serious_cat + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_18)
result_18 <- tidy(model_18)
y_pred_18 <- predict.lm(model_18, test)
y_true_18 <- test$dispo_time_filing

r_sq_18 <- round(summary(model_18)$r.squared, 4)
mae_18 <- round(MLmetrics::MAE(y_pred_18, y_true_18), 4)
mape_18 <- round(MLmetrics::MAPE(y_pred_18, y_true_18), 4)

paste("R-Squared:", r_sq_18)
paste("MAE - Mean Absolute Error:", mae_18)
paste("MAPE - Mean Absolute Percentage Error:", mape_18)

eval_filing[18,] <- c(18, r_sq_18, mae_18, mape_18)

```

### Model 19

```{r}

model_19 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age_group + gender + crime_hist + class_a + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_19)
result_19 <- tidy(model_19)
y_pred_19 <- predict.lm(model_19, test)
y_true_19 <- test$dispo_time_filing

r_sq_19 <- round(summary(model_19)$r.squared, 4)
mae_19 <- round(MLmetrics::MAE(y_pred_19, y_true_19), 4)
mape_19 <- round(MLmetrics::MAPE(y_pred_19, y_true_19), 4)

paste("R-Squared:", r_sq_19)
paste("MAE - Mean Absolute Error:", mae_19)
paste("MAPE - Mean Absolute Percentage Error:", mape_19)

eval_filing[19,] <- c(19, r_sq_19, mae_19, mape_19)

```

### Model 20

```{r}

model_20 <- lm(
  data = train,
  formula = dispo_time_filing ~ 
    age_group + gender + crime_hist + is_felony + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_20)
result_20 <- tidy(model_20)
y_pred_20 <- predict.lm(model_20, test)
y_true_20 <- test$dispo_time_filing

r_sq_20 <- round(summary(model_20)$r.squared, 4)
mae_20 <- round(MLmetrics::MAE(y_pred_20, y_true_20), 4)
mape_20 <- round(MLmetrics::MAPE(y_pred_20, y_true_20), 4)

paste("R-Squared:", r_sq_20)
paste("MAE - Mean Absolute Error:", mae_20)
paste("MAPE - Mean Absolute Percentage Error:", mape_20)

eval_filing[20,] <- c(20, r_sq_20, mae_20, mape_20)

```

### Conclusions

```{r}
eval_filing
```

## Model Selection - based on arraignment date

Initialize data frame to record key metrics

```{r}
eval_arraign <- data.frame(model_nbr = numeric(), r_squared = numeric(), mae = numeric(), mape = numeric())
```

### Model 1

```{r}

model_1 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age + age_below_18 + age_below_60 + gender + crime_hist + charge_class + total_dfa + is_trial_set + is_competency_hearing + is_drug_court + is_drug_charge
)
summary(model_1)
result_1 <- tidy(model_1)
y_pred_1 <- predict.lm(model_1, test)
y_true_1 <- test$dispo_time_arraign

r_sq_1 <- round(summary(model_1)$r.squared, 4)
mae_1 <- round(MLmetrics::MAE(y_pred_1, y_true_1), 4)
mape_1 <- round(MLmetrics::MAPE(y_pred_1, y_true_1), 4)

paste("R-Squared:", r_sq_1)
paste("MAE - Mean Absolute Error:", mae_1)
paste("MAPE - Mean Absolute Percentage Error:", mape_1)

eval_arraign[1,] <- c(1, r_sq_1, mae_1, mape_1)

```

### Model 2

```{r}

model_2 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age + gender + crime_hist + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_2)
result_2 <- tidy(model_2)
y_pred_2 <- predict.lm(model_2, test)
y_true_2 <- test$dispo_time_arraign

r_sq_2 <- round(summary(model_2)$r.squared, 4)
mae_2 <- round(MLmetrics::MAE(y_pred_2, y_true_2), 4)
mape_2 <- round(MLmetrics::MAPE(y_pred_2, y_true_2), 4)

paste("R-Squared:", r_sq_2)
paste("MAE - Mean Absolute Error:", mae_2)
paste("MAPE - Mean Absolute Percentage Error:", mape_2)

eval_arraign[2,] <- c(2, r_sq_2, mae_2, mape_2)

```

### Model 3

```{r}

model_3 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age + age_below_18 + age_below_60 + gender + crime_hist + serious_cat + total_dfa + is_trial_set + is_competency_hearing + is_drug_court + is_drug_charge
)
summary(model_3)
result_3 <- tidy(model_3)
y_pred_3 <- predict.lm(model_3, test)
y_true_3 <- test$dispo_time_arraign

r_sq_3 <- round(summary(model_3)$r.squared, 4)
mae_3 <- round(MLmetrics::MAE(y_pred_3, y_true_3), 4)
mape_3 <- round(MLmetrics::MAPE(y_pred_3, y_true_3), 4)

paste("R-Squared:", r_sq_3)
paste("MAE - Mean Absolute Error:", mae_3)
paste("MAPE - Mean Absolute Percentage Error:", mape_3)

eval_arraign[3,] <- c(3, r_sq_3, mae_3, mape_3)

```

### Model 4

```{r}

model_4 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age + gender + crime_hist + class_a + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_4)
result_4 <- tidy(model_4)
y_pred_4 <- predict.lm(model_4, test)
y_true_4 <- test$dispo_time_arraign

r_sq_4 <- round(summary(model_4)$r.squared, 4)
mae_4 <- round(MLmetrics::MAE(y_pred_4, y_true_4), 4)
mape_4 <- round(MLmetrics::MAPE(y_pred_4, y_true_4), 4)

paste("R-Squared:", r_sq_4)
paste("MAE - Mean Absolute Error:", mae_4)
paste("MAPE - Mean Absolute Percentage Error:", mape_4)

eval_arraign[4,] <- c(4, r_sq_4, mae_4, mape_4)

```

### Model 5

```{r}

model_5 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age + gender + crime_hist + is_felony + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_5)
result_5 <- tidy(model_5)
y_pred_5 <- predict.lm(model_5, test)
y_true_5 <- test$dispo_time_arraign

r_sq_5 <- round(summary(model_5)$r.squared, 4)
mae_5 <- round(MLmetrics::MAE(y_pred_5, y_true_5), 4)
mape_5 <- round(MLmetrics::MAPE(y_pred_5, y_true_5), 4)

paste("R-Squared:", r_sq_5)
paste("MAE - Mean Absolute Error:", mae_5)
paste("MAPE - Mean Absolute Percentage Error:", mape_5)

eval_arraign[5,] <- c(5, r_sq_5, mae_5, mape_5)

```

### Model 6

```{r}

model_6 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age_group + gender + crime_hist + charge_class + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_6)
result_6 <- tidy(model_6)
y_pred_6 <- predict.lm(model_6, test)
y_true_6 <- test$dispo_time_arraign

r_sq_6 <- round(summary(model_6)$r.squared, 4)
mae_6 <- round(MLmetrics::MAE(y_pred_6, y_true_6), 4)
mape_6 <- round(MLmetrics::MAPE(y_pred_6, y_true_6), 4)

paste("R-Squared:", r_sq_6)
paste("MAE - Mean Absolute Error:", mae_6)
paste("MAPE - Mean Absolute Percentage Error:", mape_6)

eval_arraign[6,] <- c(6, r_sq_6, mae_6, mape_6)

```

### Model 7

```{r}

model_7 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age_group + gender + crime_hist + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_7)
result_7 <- tidy(model_7)
y_pred_7 <- predict.lm(model_7, test)
y_true_7 <- test$dispo_time_arraign

r_sq_7 <- round(summary(model_7)$r.squared, 4)
mae_7 <- round(MLmetrics::MAE(y_pred_7, y_true_7), 4)
mape_7 <- round(MLmetrics::MAPE(y_pred_7, y_true_7), 4)

paste("R-Squared:", r_sq_7)
paste("MAE - Mean Absolute Error:", mae_7)
paste("MAPE - Mean Absolute Percentage Error:", mape_7)

eval_arraign[7,] <- c(7, r_sq_7, mae_7, mape_7)

```

### Model 8

```{r}

model_8 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age_group + gender + crime_hist + serious_cat + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_8)
result_8 <- tidy(model_8)
y_pred_8 <- predict.lm(model_8, test)
y_true_8 <- test$dispo_time_arraign

r_sq_8 <- round(summary(model_8)$r.squared, 4)
mae_8 <- round(MLmetrics::MAE(y_pred_8, y_true_8), 4)
mape_8 <- round(MLmetrics::MAPE(y_pred_8, y_true_8), 4)

paste("R-Squared:", r_sq_8)
paste("MAE - Mean Absolute Error:", mae_8)
paste("MAPE - Mean Absolute Percentage Error:", mape_8)

eval_arraign[8,] <- c(8, r_sq_8, mae_8, mape_8)

```

### Model 9

```{r}

model_9 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age_group + gender + crime_hist + class_a + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_9)
result_9 <- tidy(model_9)
y_pred_9 <- predict.lm(model_9, test)
y_true_9 <- test$dispo_time_arraign

r_sq_9 <- round(summary(model_9)$r.squared, 4)
mae_9 <- round(MLmetrics::MAE(y_pred_9, y_true_9), 4)
mape_9 <- round(MLmetrics::MAPE(y_pred_9, y_true_9), 4)

paste("R-Squared:", r_sq_9)
paste("MAE - Mean Absolute Error:", mae_9)
paste("MAPE - Mean Absolute Percentage Error:", mape_9)

eval_arraign[9,] <- c(9, r_sq_9, mae_9, mape_9)

```

### Model 10

```{r}

model_10 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age_group + gender + crime_hist + is_felony + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_court
)
summary(model_10)
result_10 <- tidy(model_10)
y_pred_10 <- predict.lm(model_10, test)
y_true_10 <- test$dispo_time_arraign

r_sq_10 <- round(summary(model_10)$r.squared, 4)
mae_10 <- round(MLmetrics::MAE(y_pred_10, y_true_10), 4)
mape_10 <- round(MLmetrics::MAPE(y_pred_10, y_true_10), 4)

paste("R-Squared:", r_sq_10)
paste("MAE - Mean Absolute Error:", mae_10)
paste("MAPE - Mean Absolute Percentage Error:", mape_10)

eval_arraign[10,] <- c(10, r_sq_10, mae_10, mape_10)

```

### Model 11

```{r}

model_11 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age + gender + crime_hist + charge_class + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_11)
result_11 <- tidy(model_11)
y_pred_11 <- predict.lm(model_11, test)
y_true_11 <- test$dispo_time_arraign

r_sq_11 <- round(summary(model_11)$r.squared, 4)
mae_11 <- round(MLmetrics::MAE(y_pred_11, y_true_11), 4)
mape_11 <- round(MLmetrics::MAPE(y_pred_11, y_true_11), 4)

paste("R-Squared:", r_sq_11)
paste("MAE - Mean Absolute Error:", mae_11)
paste("MAPE - Mean Absolute Percentage Error:", mape_11)

eval_arraign[11,] <- c(11, r_sq_11, mae_11, mape_11)

```

### Model 12

```{r}

model_12 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age + gender + crime_hist + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_12)
result_12 <- tidy(model_12)
y_pred_12 <- predict.lm(model_12, test)
y_true_12 <- test$dispo_time_arraign

r_sq_12 <- round(summary(model_12)$r.squared, 4)
mae_12 <- round(MLmetrics::MAE(y_pred_12, y_true_12), 4)
mape_12 <- round(MLmetrics::MAPE(y_pred_12, y_true_12), 4)

paste("R-Squared:", r_sq_12)
paste("MAE - Mean Absolute Error:", mae_12)
paste("MAPE - Mean Absolute Percentage Error:", mape_12)

eval_arraign[12,] <- c(12, r_sq_12, mae_12, mape_12)

```

### Model 13

```{r}

model_13 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age + gender + crime_hist + serious_cat + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_13)
result_13 <- tidy(model_13)
y_pred_13 <- predict.lm(model_13, test)
y_true_13 <- test$dispo_time_arraign

r_sq_13 <- round(summary(model_13)$r.squared, 4)
mae_13 <- round(MLmetrics::MAE(y_pred_13, y_true_13), 4)
mape_13 <- round(MLmetrics::MAPE(y_pred_13, y_true_13), 4)

paste("R-Squared:", r_sq_13)
paste("MAE - Mean Absolute Error:", mae_13)
paste("MAPE - Mean Absolute Percentage Error:", mape_13)

eval_arraign[13,] <- c(13, r_sq_13, mae_13, mape_13)

```

### Model 14

```{r}

model_14 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age + gender + crime_hist + class_a + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_14)
result_14 <- tidy(model_14)
y_pred_14 <- predict.lm(model_14, test)
y_true_14 <- test$dispo_time_arraign

r_sq_14 <- round(summary(model_14)$r.squared, 4)
mae_14 <- round(MLmetrics::MAE(y_pred_14, y_true_14), 4)
mape_14 <- round(MLmetrics::MAPE(y_pred_14, y_true_14), 4)

paste("R-Squared:", r_sq_14)
paste("MAE - Mean Absolute Error:", mae_14)
paste("MAPE - Mean Absolute Percentage Error:", mape_14)

eval_arraign[14,] <- c(14, r_sq_14, mae_14, mape_14)

```

### Model 15

```{r}

model_15 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age + gender + crime_hist + is_felony + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_15)
result_15 <- tidy(model_15)
y_pred_15 <- predict.lm(model_15, test)
y_true_15 <- test$dispo_time_arraign

r_sq_15 <- round(summary(model_15)$r.squared, 4)
mae_15 <- round(MLmetrics::MAE(y_pred_15, y_true_15), 4)
mape_15 <- round(MLmetrics::MAPE(y_pred_15, y_true_15), 4)

paste("R-Squared:", r_sq_15)
paste("MAE - Mean Absolute Error:", mae_15)
paste("MAPE - Mean Absolute Percentage Error:", mape_15)

eval_arraign[15,] <- c(15, r_sq_15, mae_15, mape_15)

```

### Model 16

```{r}

model_16 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age_group + gender + crime_hist + charge_class + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_16)
result_16 <- tidy(model_16)
y_pred_16 <- predict.lm(model_16, test)
y_true_16 <- test$dispo_time_arraign

r_sq_16 <- round(summary(model_16)$r.squared, 4)
mae_16 <- round(MLmetrics::MAE(y_pred_16, y_true_16), 4)
mape_16 <- round(MLmetrics::MAPE(y_pred_16, y_true_16), 4)

paste("R-Squared:", r_sq_16)
paste("MAE - Mean Absolute Error:", mae_16)
paste("MAPE - Mean Absolute Percentage Error:", mape_16)

eval_arraign[16,] <- c(16, r_sq_16, mae_16, mape_16)

```

### Model 17

```{r}

model_17 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age_group + gender + crime_hist + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_17)
result_17 <- tidy(model_17)
y_pred_17 <- predict.lm(model_17, test)
y_true_17 <- test$dispo_time_arraign

r_sq_17 <- round(summary(model_17)$r.squared, 4)
mae_17 <- round(MLmetrics::MAE(y_pred_17, y_true_17), 4)
mape_17 <- round(MLmetrics::MAPE(y_pred_17, y_true_17), 4)

paste("R-Squared:", r_sq_17)
paste("MAE - Mean Absolute Error:", mae_17)
paste("MAPE - Mean Absolute Percentage Error:", mape_17)

eval_arraign[17,] <- c(17, r_sq_17, mae_17, mape_17)

```

### Model 18

```{r}

model_18 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age_group + gender + crime_hist + serious_cat + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_18)
result_18 <- tidy(model_18)
y_pred_18 <- predict.lm(model_18, test)
y_true_18 <- test$dispo_time_arraign

r_sq_18 <- round(summary(model_18)$r.squared, 4)
mae_18 <- round(MLmetrics::MAE(y_pred_18, y_true_18), 4)
mape_18 <- round(MLmetrics::MAPE(y_pred_18, y_true_18), 4)

paste("R-Squared:", r_sq_18)
paste("MAE - Mean Absolute Error:", mae_18)
paste("MAPE - Mean Absolute Percentage Error:", mape_18)

eval_arraign[18,] <- c(18, r_sq_18, mae_18, mape_18)

```

### Model 19

```{r}

model_19 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age_group + gender + crime_hist + class_a + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_19)
result_19 <- tidy(model_19)
y_pred_19 <- predict.lm(model_19, test)
y_true_19 <- test$dispo_time_arraign

r_sq_19 <- round(summary(model_19)$r.squared, 4)
mae_19 <- round(MLmetrics::MAE(y_pred_19, y_true_19), 4)
mape_19 <- round(MLmetrics::MAPE(y_pred_19, y_true_19), 4)

paste("R-Squared:", r_sq_19)
paste("MAE - Mean Absolute Error:", mae_19)
paste("MAPE - Mean Absolute Percentage Error:", mape_19)

eval_arraign[19,] <- c(19, r_sq_19, mae_19, mape_19)

```

### Model 20

```{r}

model_20 <- lm(
  data = train,
  formula = dispo_time_arraign ~ 
    age_group + gender + crime_hist + is_felony + seriousness + total_dfa + is_trial_set + is_competency_hearing + is_drug_charge
)
summary(model_20)
result_20 <- tidy(model_20)
y_pred_20 <- predict.lm(model_20, test)
y_true_20 <- test$dispo_time_arraign

r_sq_20 <- round(summary(model_20)$r.squared, 4)
mae_20 <- round(MLmetrics::MAE(y_pred_20, y_true_20), 4)
mape_20 <- round(MLmetrics::MAPE(y_pred_20, y_true_20), 4)

paste("R-Squared:", r_sq_20)
paste("MAE - Mean Absolute Error:", mae_20)
paste("MAPE - Mean Absolute Percentage Error:", mape_20)

eval_arraign[20,] <- c(20, r_sq_20, mae_20, mape_20)

```

### Conclusions

```{r}
eval_arraign
```


