---
title: "BUAN 5510 - Capstone Project - EDA"
author: "Keith Castelino and Sourabh Gupta"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
	fig.width = 8.5,
	# include = FALSE,
	# echo = FALSE,
	message = FALSE,
	warning = FALSE
)

options(tibble.print_max = 40, tibble.print_min = 25)

```

# Prepare your environment

Following needs to be done in order to prepare your environment:

+ Clean environment of all variables, functions and packages
+ Load packages required for analysis (only those that are required)
+ Set colour preferences for graphs
+ Load custom functions

Read in the raw data, as is required.

```{r echo = FALSE, message = FALSE, warning = FALSE, include = FALSE}

# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environment of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE) {
  lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), 
         detach, character.only = TRUE, unload = TRUE)
}

# load required libraries (Use install.packages(tidyverse), for example, if needed)
library(tidyverse)
library(here)
library(readxl)
library(dlookr)
library(janitor)
library(stringr)
library(lubridate)
library(naniar)
library(zoo)
```

```{r}
# # LOAD IN DATA
# 
# # Data for SU Part 1 - Filed Cases and Hearings
# D11_filed_cases <- read_excel(here::here("data", "raw", "Data for SU Part 1 - Filed Cases and Hearings.xlsx"), sheet = "FiledCases SU") %>% clean_names()
# D12_case_event <- read_excel(here::here("data", "raw", "Data for SU Part 1 - Filed Cases and Hearings.xlsx"), sheet = "CaseEvents SU") %>% clean_names()
# D13_case_event_key <- read_excel(here::here("data", "raw", "Data for SU Part 1 - Filed Cases and Hearings.xlsx"), sheet = "CaseEvent Key") %>% clean_names()
# D14_case_type <- read_excel(here::here("data", "raw", "Data for SU Part 1 - Filed Cases and Hearings.xlsx"), sheet = "CaseType Key") %>% clean_names()
# 
# # Data for SU Part 2
# D21_charge <- read_excel(here::here("data", "raw", "Data for SU Part 2.xlsx"), sheet = "SU Charges") %>% clean_names()
# D22_charge_def <- read_excel(here::here("data", "raw", "Data for SU Part 2.xlsx"), sheet = "Charge Definitions") %>% clean_names()
# D23_dispo_key <- read_excel(here::here("data", "raw", "Data for SU Part 2.xlsx"), sheet = "Disposition Key") %>% clean_names()
# D24_crime_hist <- read_excel(here::here("data", "raw", "Data for SU Part 2.xlsx"), sheet = "SU CrimHist") %>% clean_names()
# 
# # Data for SU Part 3 - custody history
# D31_custody_hist <- read_excel(here::here("data", "raw", "Data for SU Part 3 - custody history.xlsx"), sheet = "CustodyHistory SU") %>% clean_names()

```

```{r message = FALSE, warning=FALSE}
#loading data
#
# xlsx files -- Data for SU Part 1 - Filed Cases and Hearings
D11_filed_cases <- read_excel("../Data/Data for SU Part 1 - Filed Cases and Hearings.xlsx", sheet = "FiledCases SU") %>% clean_names()
D12_case_event <- read_excel("../Data/Data for SU Part 1 - Filed Cases and Hearings.xlsx", sheet = "CaseEvents SU") %>% clean_names()
D13_case_event_key <- read_excel("../Data/Data for SU Part 1 - Filed Cases and Hearings.xlsx", sheet = "CaseEvent Key") %>% clean_names()
D14_case_type <- read_excel("../Data/Data for SU Part 1 - Filed Cases and Hearings.xlsx", sheet = "CaseType Key") %>% clean_names()
#
# xlsx files -- Data for SU Part 2
D21_charge <- read_excel("../Data/Data for SU Part 2.xlsx", sheet = "SU Charges") %>% clean_names()
D22_charge_def <- read_excel("../Data/Data for SU Part 2.xlsx", sheet = "Charge Definitions") %>% clean_names()
D23_dispo_key <- read_excel("../Data/Data for SU Part 2.xlsx", sheet = "Disposition Key") %>% clean_names()
D24_crime_hist <- read_excel("../Data/Data for SU Part 2.xlsx", sheet = "SU CrimHist") %>% clean_names()
#
# xlsx files -- Data for SU Part 3 - custody history
D31_custody_hist <- read_excel("../Data/Data for SU Part 3 - custody history.xlsx", sheet = "CustodyHistory SU") %>% clean_names()
#
```

# Cleaning the Data

To ensure that data is fit for analysis, we need to do the following:

+ Data structure: Ensure that the columns within each table are defined correctly
    + Confirm the data type is per expectation, and matches with definition in data dictionary
+ Derived variables: Define new columns based on data provided that are important factors for analysis

Notes for each of the datasets are below:

## D11_filed_cases

+ One record = one case
    + If there is more than one person involved in the same crime, they have different case numbers
    + A single defendant may be involved in more than one case
+ Date of Birth: Defendants need to be placed into age groups using **which date**

```{r}
D11_filed_cases <- D11_filed_cases %>% 
  mutate(
    dob_anon = ymd(dob_anon),
    event_enter_date = mdy(event_enter_date),
    age = interval(dob_anon, event_enter_date) %/% years(1),
    age_group = cut(
      age, breaks = c(0, 20, 30, 40, 50, 60, 80),
      labels = c("<=20", "20-30", "30-40", "40-50", "50-60", ">=60")
    ),
    gender = factor(gender),
    venue = factor(venue),
    venue = fct_collapse(venue, Other = c("DCS", "DCW", "JUV"))
  ) %>% 
  # select(-c(dob_anon, referral_date, event_code, police_agency, age)) %>%     #--- (-) SG
  select(-c(dob_anon, referral_date, event_code, event_enter_date,              #--- (+) SG
            police_agency, anon_le_number, age)) %>%                            #--- (+) SG

  replace_na(list(gender = "Unknown"))
```

## D12_case_event

+ One record = one event within a case
    + There is no **level** to this data
+ Event Docket Date: indicates the date of a future hearing
+ Event Code: indicates if the defendant failed to appear
    + missing values may include some DEFFTA
+ Defendant event Status contains the current status of defendant -> In jail or OUT of jain
    + need to crosscheck with custody history to fill up missing values

```{r}
D12_case_event <- D12_case_event %>% 
  mutate(event_docket_date = ymd(event_docket_date)) %>% 
  mutate_at(c("defendant_event_status", "event_code"), factor) %>% 
  select(-c(case_types, event_enter_date, venue))
```

## D13_case_event_key

We need to classify the stages of the case as follows:
+ Arraignment
+ Case Setting
+ Drug Court

```{r}
D13_case_event_key <- D13_case_event_key %>%
  mutate_at(c("hearing_code"), factor) %>% 
  mutate(
    case_setting_stage = fct_collapse(
      hearing_code,
      arraignment = c("HRARRSC"),                                                      #--- (+) SG
      case_setting = c("HRCS", "HRCSGP", "HRCSNWK", "HRCSNWS", "HRCSWK", "HRCSWS"),
      # before = c(
      #   "CSHNOWV", "DEFBWO", "DRGTH", "HR1AP", "HR1APB", 
      #   "HR2AP", "HRBOND", "HRCOMP", "HRCOMPAR",
      #   "HRCOMPBS", "HRCOMPCS", "HRCOMPKA", "HRCOMPKC", 
      #   "HRCOMPKT", "HRCOMPTR", "HRCPLSNT", "HRCRMO"
      # ),
      # after = c(
      #   "HRCTCOMP", "HROM", "HROMCON", "HRPLEA", "HRPT", "HRSNTSC", "HRSTATCF", 
      #   "LECSREC", "MOCCAP", "TRSTSC", "TRSTSCCN", "WARARRDC", "WARARRSC"
      # ),
      drug_court = c("HRARDR", "HRBWSC", "HRDR1", "HRDR2", "HRDRC")                    #--- (+) SG: others -> drug_court
    ))
  # ) %>% 
  # replace_with_na(list(case_setting_stage = c("DEFFTA", "DEFFTABT")))
```

## D14_case_type

This dataset does not map to anything, so we will ignore it. 

## D21_charge

+ Offense end date treated the same as start date if not missing
+ Additional columns:
    + we have defined if charges are enhanced

```{r}
D21_charge <- D21_charge %>% 
  mutate_at(c("offense_start_date", "offense_end_date", "filing_date", "disposition_date"), mdy) %>% 
  # mutate_at(c("disposition_code"), factor) %>% 
  mutate(
    offense_end_date = if_else(is.na(offense_end_date), offense_start_date, offense_end_date),
    charge_enhanced = if_else(is.na(enhancements), "No", "Yes"),
    charge_code = str_to_upper(str_trim(
      str_split(string = current_charge_s, pattern = '-', n = 2, simplify = TRUE)[,1]
    ))
  ) %>% 
  select(-c(referral_date, referral_charge_s, original_charge_s, warning, current_charge_s, description, enhancements))
```

## D22_charge_def

+ Charges classed as:
    + felony v/s misdemeanour
    + violent v/s non-violent

```{r}
D22_charge_def <- D22_charge_def %>% 
  mutate(
    code = if_else(str_detect(code, "[A-Z]"), code, str_pad(code, 5, "left", "0")),
    charge_class = case_when(
      class %in% c('A', 'B', 'C') ~ 2,                            # "Felony",
      class %in% c('U', 'GM') ~ 1,                                # "Misdemeanour",
      TRUE ~ 0                                                    # "Other"
    ),
    is_violent = if_else(violent == 1, 1, 0)                      # "Yes", "No/Maybe"
  ) %>% 
  select(-c(class, violent)) %>% 
  # mutate_at(c("charge_class", "is_violent"), factor) %>% 
  mutate(
    charge_class = factor(
      charge_class, levels = c(0, 1, 2), ordered = TRUE                      # c("Other", "Misdemeanour", "Felony")
    ),
    is_violent = factor(is_violent, levels = c(0, 1), ordered = TRUE)        # c("No/Maybe", "Yes")
  )
```

## D23_dispo_key

+ Description used to identify the verdict (this is an approximation)
+ Disposition type is recategorized

```{r}
D23_dispo_key <- D23_dispo_key %>% 
  mutate_at(c("type"), factor) %>% 
  mutate(
    type = fct_collapse(type, Other = c("Dismissed", "Decline or Other/Error")),
    verdict = case_when(
      str_detect(tolower(description), "guilty") ~ "guilty",
      str_detect(tolower(description), "acquit") ~ "acquitted",
      TRUE ~ "other"
    )
  ) %>% 
  mutate_at(c("verdict"), factor)
```

## D24_crime_hist

+ Confirmed that "Display Sequence" is based on the order of offenses
+ Historical records classified based on age group and charge class
+ In-state v/s Out-of-state history identified

```{r}

D24_crime_hist <- D24_crime_hist %>% 
  mutate(
    jurisdiction_state = str_extract(jursidiction, "\\b[A-Z]{2}"),
    age_group = if_else(startsWith(conviction_level, 'J'), "juvenile", "adult"),
    charge_class = if_else(endsWith(conviction_level, 'M'), "misdemeanour", "felony"),
    history_in_state = jurisdiction_state == 'WA'
  ) %>% 
  select(-c(jurisdiction_state, conviction_level))

# # to confirm that display sequence is simply order of offenses (highest value = most recent)
# 
# non_missing_defendants <- D24_crime_hist %>% 
#   group_by(defendant_id) %>% 
#   summarise(no_of_missing = sum(is.na(offense_date))) %>% 
#   filter(no_of_missing == 0) %>% 
#   select(defendant_id) %>% 
#   pull()
# 
# xxx <- D24_crime_hist %>% 
#   filter(defendant_id %in% non_missing_defendants) %>% 
#   group_by(defendant_id) %>% 
#   mutate(
#     offense_date = ymd(offense_date),
#     charge_order = desc(rank(offense_date))
#   )
# 
# xxx %>% filter(display_sequence == charge_order)

```

## D31_custody_hist

```{r}
D31_custody_hist <- D31_custody_hist %>% 
  mutate_at(c("status_update_user"), factor) %>% 
  mutate(status_update_user = fct_lump(status_update_user, other_level = "OTHER"))
```

## Combining datasets for analysis

+ Charges:
    + Representative charge: Sort charges based on class, seriousness and violent and rank in ascending order
    + Charge codes ranked in alphabetical order in case of tie
+ Criminal history: 
    + Level of data: one row per defendant
    + filter for records containing latest offense date for each defendant
    + Columns to include: age group, charge class and in/out of state jurisdiction for each defendant
    + After left joining on D11_filed_cases, mutate to find **time since last conviction**

```{r}

# # confirmed that all defendant id's are present in both file case and case event records
# setdiff(unique(D11_filed_cases$defendant_id), unique(D12_case_event$defendant_id))

# to find the representative charge

# charge_ranks <- D22_charge_def %>% 
#   arrange(charge_class, seriousness, is_violent, code) %>% 
#   mutate(charge_rank = 1:n())
# 
# rep_charge <- D21_charge %>% 
#   left_join(charge_ranks, by = c("charge_code" = "code")) %>% 
#   group_by(file_number) %>% 
#   mutate(max_charge_rank = max(charge_rank)) %>% 
#   filter(charge_rank == max_charge_rank) %>% 
#   select(file_number, charge_code)

```

```{r}
# join D21_charge and D22_charge_def to get representative charge
D21_D22_charges <- left_join(D21_charge, D22_charge_def, by = c("charge_code" = "code"))

# arrange/sort records as File_number-ASC, Class-DESC, Seriousness-DESC, Violent-DESC
D21_D22_charges <- arrange(D21_D22_charges, file_number, desc(charge_class), desc(seriousness), desc(is_violent))

# Representative charge for each file number, fetch first record of each file number
represent_charge <- D21_D22_charges[!duplicated(D21_D22_charges$file_number),] %>%
                  select(-c(offense_start_date, offense_end_date, filing_date, disposition_date,
                            disposition_code, charge_enhanced))

# this representative charges is merged with D11, filed cases
D11_cases_rep_charge <- merge(x = D11_filed_cases, y = represent_charge, by = "file_number", all.x=TRUE)

```

## Capstone Meeting Updates - 10/22/2019

1.	 representative charge for each File Number as one File Number may have multiple charges.
2.	 Critical stage of a case is case setting  (Arraignment, case setting, before case setting, after case setting)

3.	Last Arraignment Date

4.	Filling missing Defendant_Event_Status field in CaseEvents table 


  + representative charge for each File Number as one File Number may have multiple charges.


```{r}
ggplot(data=D11_cases_rep_charge, aes(x=seriousness)) +
geom_bar() +
geom_text(stat='count', aes(label=..count..), vjust=-0.5)

ggplot(data=D11_cases_rep_charge, aes(x=charge_class)) +
geom_bar() +
geom_text(stat='count', aes(label=..count..), vjust=-0.5)

ggplot(data=D11_cases_rep_charge, aes(x=is_violent)) +
geom_bar() +
geom_text(stat='count', aes(label=..count..), vjust=-0.5)
```


  + Critical stage of a case is case setting  (Arraignment, case setting, before case setting, after case setting)
  
```{r}
'%ni%' <- Negate('%in%')

# D12_case_events and D13_case_event_key to map the hearing codes
D12_D13_case_event_stages <- left_join(D12_case_event, D13_case_event_key, by = c("event_code" = "hearing_code"))

# D12_D13 last arraignment date
last_arraignment_date <- D12_D13_case_event_stages[D12_D13_case_event_stages$case_setting_stage == 'arraignment', ] %>%
                        select(-c(defendant_id, event_code, defendant_event_status, description)) %>%
                        group_by(file_number) %>% 
                        arrange(file_number, desc(event_docket_date)) %>%
                        slice(1) %>%
                        ungroup()
colnames(last_arraignment_date)[colnames(last_arraignment_date) == 'event_docket_date'] <- 'last_arr_date'
colnames(last_arraignment_date)[colnames(last_arraignment_date) == 'case_setting_stage'] <- 'arraignment'

# D12_D13 first case setting date
first_case_setting_date <- D12_D13_case_event_stages[D12_D13_case_event_stages$case_setting_stage == 'case_setting', ] %>%
                        select(-c(defendant_id, event_code, defendant_event_status, description)) %>%
                        group_by(file_number) %>% 
                        arrange(file_number, event_docket_date) %>%
                        slice(1) %>%
                        ungroup()
colnames(first_case_setting_date)[colnames(first_case_setting_date) == 'event_docket_date'] <- 'first_cs_date'
colnames(first_case_setting_date)[colnames(first_case_setting_date) == 'case_setting_stage'] <- 'case_setting'

# join 3 dataframes to get last_arraignment_date and first_case_setting_date for each file number
D12_D13_case_event_stages <- left_join(D12_D13_case_event_stages, last_arraignment_date, by = c("file_number"))
D12_D13_case_event_stages <- left_join(D12_D13_case_event_stages, first_case_setting_date, by = c("file_number"))

# new column with 4 categories: A(Arraignment), B(Before Case Setting), C(Case Setting), D(After Case Setting)
D12_D13_case_event_stages <-D12_D13_case_event_stages %>%  
      mutate(case_stage = ifelse(event_docket_date == last_arr_date, 'A(Arraignment)', 
                                ifelse(event_docket_date > last_arr_date & event_docket_date < first_cs_date, 'B(Before Case Setting)',
                                  ifelse(event_docket_date == first_cs_date, 'C(Case Setting)',
                                        ifelse(event_docket_date > first_cs_date, 'D(After Case Setting)', 'Question')))))

# arrange/sort records as File_number-ASC, Docket_date-ASC
# D12_D13_case_event_stages <- arrange(D12_D13_case_event_stages, file_number, desc(event_docket_date))

# Convert age into categories as mentioned in Data dictionary
# D12_D13_case_event_stages <-D12_D13_case_event_stages %>%  
#                             mutate(case_stage = ifelse(case_setting_stage %ni% c('arraignment', 'case_setting'), '<19', 
#                                           ifelse(age > 19 & age < 30, '20-29',
#                                                 ifelse(age >= 30 & age < 40, '30-39',
#                                                       ifelse(age >=40  & age < 50, '40-49',
#                                                             ifelse(age >= 50 & age < 60, '50-59',
#                                                                   ifelse(age >= 60, '>60',0))))))) %>%
#                   select(c(file_number, defendant_id, dob_anon, gender, age_cate))
```



