---
title: "BUAN 5510 - Capstone Project - EDA"
author: "Keith Castelino and Sourabh Gupta"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
	fig.width = 8.5,
	# include = FALSE,
	# echo = FALSE,
	message = FALSE,
	warning = FALSE
)

options(tibble.print_max = 40, tibble.print_min = 25)

```

# Prepare your environment

Following needs to be done in order to prepare your environment:

+ Clean environment of all variables, functions and packages
+ Load packages required for analysis (only those that are required)
+ Set colour preferences for graphs
+ Load custom functions

Read in the raw data, as is required.

```{r echo = FALSE, message = FALSE, warning = FALSE, include = FALSE}

# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environment of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE) {
  lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), 
         detach, character.only = TRUE, unload = TRUE)
}

# load required libraries (Use install.packages(tidyverse), for example, if needed)
library(tidyverse)
library(here)
library(readxl)
library(dlookr)
library(janitor)
library(stringr)
library(lubridate)
library(naniar)
library(zoo)
library(birk)

# LOAD IN DATA

# Data for SU Part 1 - Filed Cases and Hearings
D11_filed_cases <- read_excel(here::here("data", "raw", "Data for SU Part 1 - Filed Cases and Hearings.xlsx"), sheet = "FiledCases SU") %>% clean_names()
D12_case_event <- read_excel(here::here("data", "raw", "Data for SU Part 1 - Filed Cases and Hearings.xlsx"), sheet = "CaseEvents SU") %>% clean_names()
D13_case_event_key <- read_excel(here::here("data", "raw", "Data for SU Part 1 - Filed Cases and Hearings.xlsx"), sheet = "CaseEvent Key") %>% clean_names()
D14_case_type <- read_excel(here::here("data", "raw", "Data for SU Part 1 - Filed Cases and Hearings.xlsx"), sheet = "CaseType Key") %>% clean_names()

# Data for SU Part 2
D21_charge <- read_excel(here::here("data", "raw", "Data for SU Part 2.xlsx"), sheet = "SU Charges") %>% clean_names()
D22_charge_def <- read_excel(here::here("data", "raw", "Data for SU Part 2.xlsx"), sheet = "Charge Definitions") %>% clean_names()
D23_dispo_key <- read_excel(here::here("data", "raw", "Data for SU Part 2.xlsx"), sheet = "Disposition Key") %>% clean_names()
D24_crime_hist <- read_excel(here::here("data", "raw", "Data for SU Part 2.xlsx"), sheet = "SU CrimHist") %>% clean_names()

# Data for SU Part 3 - custody history
D31_custody_hist <- read_excel(here::here("data", "raw", "Data for SU Part 3 - custody history.xlsx"), sheet = "CustodyHistory SU") %>% clean_names()

```

# Cleaning the Data

To ensure that data is fit for analysis, we need to do the following:

+ Data structure: Ensure that the columns within each table are defined correctly
    + Confirm the data type is per expectation, and matches with definition in data dictionary
+ Derived variables: Define new columns based on data provided that are important factors for analysis

Notes for each of the datasets are below:

## D11_filed_cases

+ One record = one case
    + If there is more than one person involved in the same crime, they have different case numbers
    + A single defendant may be involved in more than one case
+ Date of Birth: Defendants need to be placed into age groups using **which date**

```{r}
D11_filed_cases <- D11_filed_cases %>% 
  mutate(
    dob_anon = ymd(dob_anon),
    event_enter_date = mdy(event_enter_date),
    age = interval(dob_anon, event_enter_date) %/% years(1),
    age_group = cut(
      age, breaks = c(0, 18, 30, 40, 50, 60, 80),
      labels = c("<=18", "20-30", "30-40", "40-50", "50-60", ">=60")
    ),
    venue = fct_collapse(venue, Other = c("DCS", "DCW", "JUV"))
  ) %>% 
  mutate_at(c("gender", "venue", "case_types"), factor) %>% 
  select(-c(dob_anon, referral_date, event_code, police_agency, age, anon_le_number))
```

## D12_case_event

+ One record = one event within a case
    + There is no **level** to this data
+ Event Docket Date: indicates the date of a future hearing
+ Event Code: indicates if the defendant failed to appear
    + missing values may include some DEFFTA
    + need to crosscheck with custody history to fill up missing values

```{r}
D12_case_event <- D12_case_event %>% 
  mutate(event_docket_date = ymd(event_docket_date)) %>% 
  mutate_at(c("defendant_event_status", "event_code"), factor) %>% 
  select(-c(case_types, event_enter_date, venue))
```

## D13_case_event_key

+ Case stages need to be classified in 4 ways:
    + Arraignment
    + Pre Case Setting
    + Case Setting
    + Post Case Setting
+ We know what the following codes mean:
    + HRARRSC: Arraignment
    + HRCSGP, HRCSNWK, HRCSNWS, HRCSWK, HRCSWS: Case setting
+ Pre Case Setting: Any event that is after the last arraignment date
+ Post Case Setting: Any event that is after the first case setting date

```{r}

# D13_case_event_key <- D13_case_event_key %>%
#   mutate_at(c("hearing_code"), factor) %>% 
#   mutate(
#     case_setting_stage = fct_collapse(
#       hearing_code,
#       during = c("HRCS", "HRCSGP", "HRCSNWK", "HRCSNWS", "HRCSWK", "HRCSWS"),
#       before = c(
#         "CSHNOWV", "DEFBWO", "DRGTH", "HR1AP", "HR1APB", 
#         "HR2AP", "HRARRSC", "HRBOND", "HRCOMP", "HRCOMPAR",
#         "HRCOMPBS", "HRCOMPCS", "HRCOMPKA", "HRCOMPKC", 
#         "HRCOMPKT", "HRCOMPTR", "HRCPLSNT", "HRCRMO"
#       ),
#       after = c(
#         "HRCTCOMP", "HROM", "HROMCON", "HRPLEA", "HRPT", "HRSNTSC", "HRSTATCF", 
#         "LECSREC", "MOCCAP", "TRSTSC", "TRSTSCCN", "WARARRDC", "WARARRSC"
#       ),
#       other = c("HRARDR", "HRBWSC", "HRDR1", "HRDR2", "HRDRC")
#     )
#   ) %>% 
#   replace_with_na(list(case_setting_stage = c("DEFFTA", "DEFFTABT")))

```

## D14_case_type

This dataset does not map to anything, so we will ignore it. 

## D21_charge

+ Offense end date treated the same as start date if not missing
+ Additional columns:
    + we have defined if charges are enhanced

```{r}
D21_charge <- D21_charge %>% 
  mutate_at(c("offense_start_date", "offense_end_date", "filing_date", "disposition_date"), mdy) %>% 
  # mutate_at(c("disposition_code"), factor) %>% 
  mutate(
    offense_end_date = if_else(is.na(offense_end_date), offense_start_date, offense_end_date),
    charge_enhanced = if_else(is.na(enhancements), "No", "Yes"),
    charge_code = str_to_upper(str_trim(
      str_split(string = current_charge_s, pattern = '-', n = 2, simplify = TRUE)[,1]
    ))
  ) %>% 
  select(-c(referral_date, referral_charge_s, original_charge_s, warning, current_charge_s, description, enhancements))
```

## D22_charge_def

+ Charges classed as:
    + felony v/s misdemeanour
    + violent v/s non-violent

```{r}
D22_charge_def <- D22_charge_def %>% 
  mutate(
    code = if_else(str_detect(code, "[A-Z]"), code, str_pad(code, 5, "left", "0")),
    charge_class = case_when(
      class %in% c('A', 'B', 'C') ~ "Felony",
      class %in% c('U', 'GM') ~ "Misdemeanour",
      TRUE ~ "Other"
    ),
    is_violent = if_else(violent == 1, "Yes", "No/Maybe")
  ) %>% 
  select(-c(violent)) %>% 
  # mutate_at(c("charge_class", "is_violent"), factor) %>% 
  mutate(
    charge_class = factor(
      charge_class, levels = c("Other", "Misdemeanour", "Felony"), ordered = TRUE
    ),
    is_violent = factor(is_violent, levels = c("No/Maybe", "Yes"), ordered = TRUE)
  )
```

## D23_dispo_key

+ Description used to identify the verdict (this is an approximation)
+ Disposition type is recategorized

```{r}
D23_dispo_key <- D23_dispo_key %>% 
  mutate_at(c("type"), factor) %>% 
  mutate(
    type = fct_collapse(type, Other = c("Dismissed", "Decline or Other/Error")),
    verdict = case_when(
      str_detect(tolower(description), "guilty") ~ "guilty",
      str_detect(tolower(description), "acquit") ~ "acquitted",
      TRUE ~ "other"
    )
  ) %>% 
  mutate_at(c("verdict"), factor)
```

## D24_crime_hist

+ Confirmed that "Display Sequence" is based on the order of offenses
+ Historical records classified based on age group and charge class
+ In-state v/s Out-of-state history identified

```{r}

D24_crime_hist <- D24_crime_hist %>% 
  mutate(
    jurisdiction_state = str_extract(jursidiction, "\\b[A-Z]{2}"),
    age_group = if_else(startsWith(conviction_level, 'J'), "juvenile", "adult"),
    charge_class = if_else(endsWith(conviction_level, 'M'), "misdemeanour", "felony"),
    history_in_state = jurisdiction_state == 'WA'
  ) %>% 
  select(-c(jurisdiction_state, conviction_level))

# # to confirm that display sequence is simply order of offenses (highest value = most recent)
# 
# non_missing_defendants <- D24_crime_hist %>% 
#   group_by(defendant_id) %>% 
#   summarise(no_of_missing = sum(is.na(offense_date))) %>% 
#   filter(no_of_missing == 0) %>% 
#   select(defendant_id) %>% 
#   pull()
# 
# xxx <- D24_crime_hist %>% 
#   filter(defendant_id %in% non_missing_defendants) %>% 
#   group_by(defendant_id) %>% 
#   mutate(
#     offense_date = ymd(offense_date),
#     charge_order = desc(rank(offense_date))
#   )
# 
# xxx %>% filter(display_sequence == charge_order)

```

## D31_custody_hist

```{r}

D31_custody_hist <- D31_custody_hist %>% 
  mutate_at(c("current_status", "custody_status_history", "status_update_user"), factor) %>% 
  mutate(
    status_update_user = fct_lump(status_update_user, other_level = "OTHER"),
    status_history_date = excel_numeric_to_date(as.numeric(status_history_date), "modern"),
    current_status_date = as.Date(current_status_date)
  )

```

## Data Quality Issues

+ D21_charge:
    + There are 36 records where the filing date is earlier than the earliest offense start date
    + There are 1731 records where the filing date is earlier than the offense start date

```{r}

# D21_charge %>% 
#   group_by(file_number) %>% 
#   mutate(min_offense_start = min(offense_start_date)) %>% 
#   filter(filing_date < min_offense_start)
# 
# D21_charge %>% 
#   filter(filing_date < offense_start_date)

```


## Further data cleaning

The data is further trimmed based on the following:

+ D11_filed_cases:
    + Case types that occur less than 1% are grouped as "Other"
    + Case types with missing values are classified as "Unknown"
    + Missing values replaced in age group and gender variables
+ D12_case_event:
    + Case events with missing docket dates are eliminated
    + CSHNOWV: will never have a docket date
    + HRCS, TRSTSCCN, HRCSGP, HROMCON, HRSNTSC: missing due to (possibly) human error
    + most other event codes with missing values each occur in less than 1% of events
+ D31_custody_history:
    + Filling 

```{r}

# classifying case types that occur less than 1% as "Other"
# missing case types classified as "Unknown"

# D11_filed_cases %>%
#   diagnose_category(top = 25, add_character = TRUE) %>%
#   filter(variables == "case_types") %>%
#   arrange(desc(freq))

D11_filed_cases <- D11_filed_cases %>%
  mutate(
    # case_types = fct_lump(case_types, prop = 0.01),
    case_types = fct_explicit_na(case_types, na_level = "Unknown"),
    age_group = fct_explicit_na(age_group, na_level = "Unknown")
  ) %>% 
  replace_na(list(gender = "Unknown"))

D12_case_event <- D12_case_event %>% 
  filter(!(is.na(event_docket_date) & event_code != "DEFFTA"))

D21_charge <- D21_charge %>% 
  filter(!(is.na(disposition_date) | is.na(disposition_code)))

```

### Missing values for defendant event status

Highlight all lines and hit Ctrl+C to uncomment

Need to read in the CSV file to work with it for further EDA

```{r}

# # condition_1 : if defendant status is not missing, use defendant_event_status
# # condition_2 : if docket_date is >= status_history_date and < current_status_date than custody_status_history
# # else: current_status
# 
# case_events <- D12_case_event %>% 
#   mutate(
#     event_docket_date_1 = if_else(
#       is.na(event_docket_date), lag(event_docket_date), event_docket_date
#     )
#   ) %>% 
#   filter(is.na(defendant_event_status))
# 
# get_near_status_history_dt <- function(def_id) {
#   D31_custody_hist$status_history_date[D31_custody_hist$defendant_id == def_id]
# }
# get_near_current_status_dt <- function(def_id) {
#   D31_custody_hist$current_status_date[D31_custody_hist$defendant_id == def_id]
# }
# 
# case_events$nearest_status_hist_date <- as.Date(NA)
# case_events$nearest_current_status_date <- as.Date(NA)
# 
# for (i in 1:nrow(case_events)) {
#   def_id <- case_events$defendant_id[i]
#   status_hist_dt <- get_near_status_history_dt(def_id)
#   current_docket_dt <- case_events$event_docket_date[i]
#   case_events$nearest_status_hist_date[i] <- as.Date(
#     status_hist_dt[which.closest(status_hist_dt, current_docket_dt)]
#   )
# }
# 
# for (i in 1:nrow(case_events)) {
#   def_id <- case_events$defendant_id[i]
#   curr_status_dt <- get_near_current_status_dt(def_id)
#   current_docket_dt <- case_events$event_docket_date[i]
#   case_events$nearest_current_status_date[i] <- as.Date(
#     curr_status_dt[which.closest(curr_status_dt, current_docket_dt)]
#   )
# }
# 
# events_def_status_na_fill <- left_join(
#   case_events, D31_custody_hist, 
#   by = c(
#     "defendant_id" = "defendant_id", 
#     "nearest_status_hist_date" = "status_history_date", 
#     "nearest_current_status_date" = "current_status_date"
#   )
# ) %>% 
#   mutate(
#     correct_defendant_status = if_else(
#       condition = !is.na(defendant_event_status),
#       true = defendant_event_status,
#       false = if_else(
#         condition = event_docket_date >= nearest_status_hist_date & 
#           event_docket_date < nearest_current_status_date,
#         true = custody_status_history,
#         false = current_status
#       )
#     )
#   ) %>% 
#   select(
#     file_number, defendant_id, event_code, event_docket_date, correct_defendant_status
#   ) %>% 
#   rename(defendant_status = correct_defendant_status)
# 
# events <- D12_case_event %>% 
#   filter(!is.na(defendant_event_status)) %>% 
#   select(file_number, defendant_id, event_code, event_docket_date, defendant_event_status) %>% 
#   rename(defendant_status = defendant_event_status)
# 
# events_with_missing_status <- bind_rows(events_def_status_na_fill, events)
# 
# write_csv(
#   events_with_missing_status, 
#   path = here::here("data", "processed", "events_with_missing_def_status.csv")
# )

```

### Time to disposition

Time to disposition for a case is measured in "days" and is the difference between the earliest filing date and the latest disposition date.

```{r}

time_to_disposition <- D21_charge %>% 
  group_by(file_number) %>% 
  mutate(
    earliest_filing_date = min(filing_date),
    latest_disposition_date = max(disposition_date),
    time_to_disposition_days = difftime(
      earliest_filing_date, latest_disposition_date, "days"
    )
  ) %>% 
  select(-c("earliest_filing_date", "latest_disposition_date"))

```

### Finding the representative charge:

+ Representative charge: Sort charges based on class, seriousness and violent and rank in ascending order
+ Charge codes ranked in alphabetical order in case of tie
+ Duplicate charge codes are eliminated by filtering out instances of charge codes after the first instance

```{r}

# # confirmed that all defendant id's are present in both file case and case event records
# setdiff(unique(D11_filed_cases$defendant_id), unique(D12_case_event$defendant_id))

# to find the representative charge

charge_code_unique <- D22_charge_def %>% 
  arrange(desc(charge_class), desc(seriousness), desc(is_violent), code) %>% 
  mutate(charge_rank = 1:n()) %>% 
  group_by(code) %>% 
  arrange(desc(charge_rank)) %>% 
  mutate(charge_code_instance = 1:n()) %>% 
  filter(charge_code_instance == 1) %>% 
  select(-charge_code_instance)

rep_charge <- D21_charge %>% 
  left_join(charge_code_unique, by = c("charge_code" = "code")) %>% 
  group_by(file_number) %>% 
  select(file_number, charge_code, charge_rank) %>% 
  filter(!is.na(charge_code)) %>% 
  mutate(max_charge_rank = min(charge_rank)) %>% 
  filter(charge_rank == max_charge_rank) %>% 
  select(file_number, charge_code) %>% 
  unique()

```

### Case Setting Stages

```{r}

last_arraignment <- D12_case_event %>% 
  filter(event_code == "HRARRSC") %>% 
  group_by(file_number) %>% 
  summarise(arraignment_date = max(event_docket_date))

first_case_setting <- D12_case_event %>% 
  filter(event_code %in% c("HRCSGP", "HRCSNWK", "HRCSNWS", "HRCSWK", "HRCSWS")) %>% 
  group_by(file_number) %>% 
  summarise(case_setting_date = min(event_docket_date))

```

```{r}
D12_case_event %>%  
  mutate(
    event_group = fct_collapse(
      event_code,
      arraignment = c("HRARRSC"),
      case_setting = c("HRCSGP", "HRCSNWK", "HRCSNWS", "HRCSWK", "HRCSWS"), 
      group_other = TRUE
    )
  ) %>% 
  group_by(event_group) %>% 
  summarise(no_of_cases = n_distinct(file_number)) %>% 
  arrange(desc(no_of_cases))
```

