---
title: "BUAN 5510 - Capstone Project - Linear Regression"
author: "Keith Castelino and Sourabh Gupta"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
	fig.width = 8.5,
	# include = FALSE,
	# echo = FALSE,
	message = FALSE,
	warning = FALSE
)

options(tibble.print_max = 40, tibble.print_min = 25)

library(tidyverse)
library(here)
library(broom)
library(caTools)
library(MLmetrics)
library(MASS)
library(tree)
library(fastDummies)

```

# Read in the Data

Read in the dataset.

```{r}

dispo_by_category <- read_csv(here::here("data", "processed", "dispo_by_category.csv"))
represent_charge_by_file <- read_csv(here::here("data", "processed", "represent_charge_by_file.csv"))
case_competency_hearing <- read_csv(here::here("data", "processed", "case_competency_hearing.csv"))
case_trial_set_event <- read_csv(here::here("data", "processed", "case_trial_set_event.csv"))
defendant_ages <- read_csv(here::here("data", "processed", "defendant_ages.csv"))
# defendant_seriousness <- read_csv(here::here("data", "processed", "defendant_seriousness.csv"))
defendant_drug_charge <- read_csv(here::here("data", "processed", "defendant_drug_charge.csv"))
defendant_drug_court <- read_csv(here::here("data", "processed", "defendant_drug_court.csv"))

base_data <- dispo_by_category %>% 
  # select(-c(age, age_group)) %>% 
  left_join(represent_charge_by_file, by = c("file_number" = "file_number")) %>% 
  left_join(case_competency_hearing, by = c("file_number" = "file_number")) %>% 
  left_join(case_trial_set_event, by = c("file_number" = "file_number")) %>% 
  # left_join(defendant_ages %>% select(-c(defendant_id, age_group)), by = c("file_number" = "file_number")) %>% 
  # left_join(defendant_seriousness %>% select(-defendant_id), by = c("file_number" = "file_number")) %>% 
  left_join(defendant_drug_charge, by = c("file_number" = "file_number")) %>% 
  left_join(defendant_drug_court, by = c("file_number" = "file_number")) %>% 
  replace_na(list(class.x = "Other", serious_cat = "Other", defendant_status = "Other", seriousness = 0, no_of_trial_sets = 0)) %>% 
  mutate(
    is_rep_charge = if_else(charge_code.x == charge_code | is.na(charge_code.x), 1, 0),
    class.x = if_else(class.x %in% c("A", "B", "C"), class.x, "GM"),
    class_a = if_else(class.x == "A", 1, 0),
    age_below_18 = if_else(age <= 18, 1, 0),
    age_below_60 = if_else(age >= 60, 1, 0),
    is_felony = if_else(class.x %in% c("A", "B", "C"), 1, 0),
    is_drug_charge = if_else(no_of_drug_charges > 0, 1, 0),
    is_drug_court = if_else(no_of_drug_court > 0, 1, 0),
    is_trial_set = if_else(no_of_trial_sets > 0, 1, 0)
  ) %>% 
  filter(
    !is.na(dispo_time_arraign),
    age_group != "Unknown",
    gender != "Unknown",
    class.x %in% c("M", "A", "B", "C", "GM"),
    defendant_status %in% c("IN", "OUT", "Other"),
    between(dispo_time_filing, 1, 5 * 365),
    between(dispo_time_arraign, 1, 5 * 365)
  ) %>% 
  mutate_at(c("age_group", "gender", "crime_hist", "class.x", "serious_cat", "defendant_status"), factor) %>% 
  dplyr::select(
    file_number, defendant_id,
    dispo_time_filing, dispo_time_arraign, gender, age, age_group, age_below_18, age_below_60,
    class.x, class_a, is_felony, seriousness, serious_cat, crime_hist, total_dfa,
    no_of_drug_charges, no_of_drug_court, no_of_trial_sets,
    is_drug_charge, is_drug_court, is_trial_set, is_competency_hearing #, defendant_status
  ) %>%
  rename(charge_class = class.x)

base_data$age_group <- relevel(base_data$age_group, "LE_18")
base_data$gender <- relevel(base_data$gender, "Female")
base_data$crime_hist <- relevel(base_data$crime_hist, "No")
base_data$charge_class <- relevel(base_data$charge_class, "GM")
# base_data$defendant_status <- relevel(base_data$defendant_status, "IN")

```

Load in the functions to display output for the decision tree

```{r}

print_tree <- function(reg_tree, tree_title = "Title required here") {
  plot(reg_tree)
  text(reg_tree, pretty = 0)
  title(main = tree_title)
}

print_tree_metrics <- function(reg_tree, test_set, y_variable) {
  
  tree_summary <- summary(reg_tree)
  
  y_pred <- predict(reg_tree, newdata = test_set)
  y_true <- test_set[ , y_variable] %>% pull()
  
  mae <- round(MLmetrics::MAE(y_pred, y_true), 4)
  mape <- round(MLmetrics::MAPE(y_pred, y_true), 4)
  
  print(tree_summary)
  
  tree_mae <- paste("MAE - Mean Absolute Error:", mae)
  tree_mape <- paste("MAPE - Mean Absolute Percentage Error:", mape)
  
  print(tree_mae)
  print(tree_mape)
  
}

print_cross_val_plot <- function(reg_tree, training_set) {
  cv_tree <- cv.tree(reg_tree)
  cv_plot <- plot(
    cv_tree$size, sqrt(cv_tree$dev / nrow(training_set)), 
    type = "b", xlab = "Tree Size", ylab = "CV-RMSE",
    main = "Cross validation plot"
  )
  print(cv_plot)
}

```

Limit the variables in the decision tree to those that are required. Hence, remove the file number values and defendant id's

Set dummy variables to take care of categorical data

```{r}

base_data <- base_data %>% dplyr::select(-c(file_number, defendant_id, dispo_time_arraign))

# set dummy variables

base_dataset <- fastDummies::dummy_cols(base_data, remove_first_dummy = TRUE) %>% 
  dplyr::select(-c(gender, age_group, charge_class, serious_cat, crime_hist)) # defendant_status

```

Split into test and train sets

```{r}
set.seed(101)
sample <- sample.split(base_dataset$dispo_time_filing, SplitRatio = 0.80 * nrow(base_dataset))
train <- subset(base_dataset, sample == TRUE)
test <- subset(base_dataset, sample == FALSE)
```

# Decision Tree and Pruning

```{r}

dispo_tree_1 <- tree(
  dispo_time_filing ~ 
    age # + age_below_18 + age_below_60
    + class_a + is_felony + seriousness + total_dfa
    + no_of_drug_charges + no_of_drug_court + no_of_trial_sets 
    + is_drug_charge + is_drug_court + is_trial_set + is_competency_hearing + gender_Female
    # + `age_group_20-30` + `age_group_30-40` + `age_group_50-60` + age_group_GE_60 + age_group_LE_18
    + charge_class_C +  charge_class_GM + charge_class_B
    # + `serious_cat_<=5` + `serious_cat_>10` + serious_cat_Other 
    + crime_hist_No
  , 
  data = train
)

dispo_tree_2 <- tree(
  dispo_time_filing ~ 
    age # + age_below_18 + age_below_60
    + class_a + is_felony + seriousness + total_dfa
    # + no_of_drug_charges + no_of_drug_court + no_of_trial_sets 
    + is_drug_charge + is_drug_court + is_trial_set + is_competency_hearing + gender_Female
    # + `age_group_20-30` + `age_group_30-40` + `age_group_50-60` + age_group_GE_60 + age_group_LE_18
    + charge_class_C +  charge_class_GM + charge_class_B
    # + `serious_cat_<=5` + `serious_cat_>10` + serious_cat_Other 
    + crime_hist_No
  , 
  data = train
)

dispo_tree_prune_5 <- prune.tree(dispo_tree_1, best = 5)
dispo_tree_prune_7 <- prune.tree(dispo_tree_1, best = 7)

```

This tree uses both no of quantitative and dummy variables for drug court cases, drug charges and trial setting events.

```{r}
print_tree_metrics(dispo_tree_1, test, "dispo_time_filing")
print_tree(dispo_tree_1, "Unpruned Tree")
print_cross_val_plot(dispo_tree_1, train)
```

This tree uses only the dummy variables relating to drug court cases, drug charges and trial setting events

```{r}
print_tree_metrics(dispo_tree_2, test, "dispo_time_filing")
print_tree(dispo_tree_2, "Unpruned Tree")
print_cross_val_plot(dispo_tree_2, train)
```

Pruning the decision tree to 5 nodes??

```{r include = FALSE}
print_tree(dispo_tree_prune_5)
print_tree_metrics(dispo_tree_prune_5, test)
```

Pruning the decision tree to 7 nodes??

```{r include = FALSE}
print_tree(dispo_tree_prune_7)
print_tree_metrics(dispo_tree_prune_7, test)
```

The cross validation plot shows that a decision tree with 8 nodes explains the relationships best. This is evident in the much higher MAE and MAPE that we find when pruning the tree to 5 or 7 nodes. Hence pruning is unnecessary (Note: output isn't shown here)